Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-02-18 manuscript/supplemental_material.Rnw
fix float; gene tables and caching in supplemental
- use float for table captions on top
- fix species table legend
- fix cache location to fix cache conflicts
- use better gene tree font sizes
- clear pages after tables

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")

@@

\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Common Name        & Reference & Read Length \\
    \midrule
    \textit{Ateles fusciceps}     & Spider monkey      & \textit{Homo sapiens} & 100 \\
    \textit{Bos taurus}           & Cow                & UMD3.1     & 75 \\
    \textit{Canis familiaris}     & Dog                & CanFam3.1  & 150 \\
    \textit{Dasypus novemcinctus} & 9-Banded Armadillo & DASNOV3    & 100 \\
    \textit{Equus caballus}       & Horse              & EquCab2    & 100 \\ 
    \textit{Loxodonta africana}   & Elephant           & loxAfr3    & 75 \\
    \textit{Homo sapiens}         & Human              & GRCh38     & 75, 150 \\
    \textit{Monodelphis domestica} & Opossum           & BROAD05    & 75 \\
    \textit{Mus musculus}         & Mouse              & GRCm38     & 35,44,150 \\
    \textit{Nannospalax galili}   & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Ovis aries}           & Sheep              & Oar\_v3.1   & 151 \\
    \textit{Pan paniscus}         & Bonobo             & CHIMP2.1.4 & 150 \\
    \textit{Spalax carmeli}       & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Sus scrofa}           & Duroc (pig)        & Sscrofa10.2 & 51 \\
    \bottomrule
  \end{tabular}
  \caption{Reference genomes for aligned species}
  \label{tab:ref_genome_aligned_species}
\end{table}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="FSTL1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,min.fpkm=5)
@ 
\caption{\genetreeexplanation{follistatin-like 1 \textit{FSTL1}}}
\label{fig:fstl1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="SERPINE2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,min.fpkm=1)
@ 
\caption{\genetreeexplanation{\textit{SERPINE2}}}
\label{fig:serpine_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_id=="ENSCAFG00000010638" & species=="canis familiaris",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{\textit{SSLP1}}}
\label{fig:sslp1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CYP19A1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{estrogen synthesis genes}}
\label{fig:estrogen_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HSD3B1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSMODG00000005385","ENSLAFG00000027896")
                 )
@ 
\caption{\genetreeexplanation{progesterone synthesis genes}}
\label{fig:progesterone_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CSH1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{chorionic somatomammotropin}}
\label{fig:csh1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{prolactin}}
\label{fig:prolactin_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="RLN2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{\textit{RLN2}}}
\label{fig:rln2_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree["ENSDNOG00000033320",tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSDNOG00000033320","ENSMODG00000017988"))
@ 
\caption{\genetreeexplanation{\cow \textit{OVOS2}}}
\label{fig:ovos2_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRLR" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSG00000113494","ENSMODG00000020376"))
@ 
\caption{\genetreeexplanation{prolactin receptor, \textit{PRLR}}}
\label{fig:prolactin_receptor_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HOXA11" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSBTAG00000014738","ENSECAG00000013229")
                 )
@ 
\caption{\genetreeexplanation{\textit{HOXA11}}}
\label{fig:hoxa11_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HBE1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{Expression of \textit{HBE1} and related genes}
\label{fig:hbe1_evolution_tree}
\end{figure}

@@

\begin{table}
  \footnotesize
<<rodentia.lineage.specific,echo=FALSE,results="asis">>=
rod.all.lin.spec <- placenta.lineage.specific.table("rod.all")
setkey(rod.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed significantly in the Rodentia
  order as compared to all other analyzed species.}
\label{tab:rodentia_lineage_specific}
\end{table}

@@

\begin{table}
  \footnotesize
<<eua.lau.lineage.specific,echo=FALSE,results="asis">>=
eua.lau.lin.spec <- placenta.lineage.specific.table("eua.lau")
setkey(eua.lau.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed between the
  Euasiatherian and Laurasiatherian superorder.}
\label{tab:laurasiatheria_eurasiatherian_lineage_specific}
\end{table}

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE,cache.path="supplemental_cache/")
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")

@@

\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Common Name        & Reference & Read Length \\
    \midrule
    \textit{Ateles fusciceps}     & Spider monkey      & \textit{Homo sapiens} & 100 \\
    \textit{Bos taurus}           & Cow                & UMD3.1     & 75 \\
    \textit{Canis familiaris}     & Dog                & CanFam3.1  & 150 \\
    \textit{Dasypus novemcinctus} & 9-Banded Armadillo & DASNOV3    & 100 \\
    \textit{Equus caballus}       & Horse              & EquCab2    & 100 \\ 
    \textit{Loxodonta africana}   & Elephant           & loxAfr3    & 75 \\
    \textit{Homo sapiens}         & Human              & GRCh38     & 75, 150 \\
    \textit{Monodelphis domestica} & Opossum           & BROAD05    & 75 \\
    \textit{Mus musculus}         & Mouse              & GRCm38     & 35,44,150 \\
    \textit{Nannospalax galili}   & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Ovis aries}           & Sheep              & Oar\_v3.1   & 151 \\
    \textit{Pan paniscus}         & Bonobo             & CHIMP2.1.4 & 150 \\
    \textit{Spalax carmeli}       & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Sus scrofa}           & Duroc (pig)        & Sscrofa10.2 & 51 \\
    \bottomrule
  \end{tabular}
  \caption{Species names, common names, reference genomes and read 
    lengths for aligned species. An italic species as a reference 
    indicates that assembled reads were annotated using the corresponding
    reference genome as explained in the text.}
  \label{tab:ref_genome_aligned_species}
\end{table}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="FSTL1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=5,
                 axis.text.size=5,
                 fontsize=8
                 )
@ 
\caption{\genetreeexplanation{follistatin-like 1 \textit{FSTL1}}}
\label{fig:fstl1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="SERPINE2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=1,
                 axis.text.size=5,
                 fontsize=4)
@ 
\caption{\genetreeexplanation{\textit{SERPINE2}}}
\label{fig:serpine_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_id=="ENSCAFG00000010638" & species=="canis familiaris",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=6)
@ 
\caption{\genetreeexplanation{\textit{SSLP1}}}
\label{fig:sslp1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CYP19A1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=7)
@ 
\caption{\genetreeexplanation{estrogen synthesis genes}}
\label{fig:estrogen_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HSD3B1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 subtree=c("ENSMODG00000005385","ENSLAFG00000027896"),
                 axis.text.size=5,
                 fontsize=6
                 )
@ 
\caption{\genetreeexplanation{progesterone synthesis genes}}
\label{fig:progesterone_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CSH1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=4)
@ 
\caption{\genetreeexplanation{chorionic somatomammotropin}}
\label{fig:csh1_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=5)
@ 
\caption{\genetreeexplanation{prolactin}}
\label{fig:prolactin_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="RLN2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=5)
@ 
\caption{\genetreeexplanation{\textit{RLN2}}}
\label{fig:rln2_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree["ENSDNOG00000033320",tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 subtree=c("ENSDNOG00000033320","ENSMODG00000017988"),
                 axis.text.size=5,
                 fontsize=5)
@ 
\caption{\genetreeexplanation{\cow \textit{OVOS2}}}
\label{fig:ovos2_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRLR" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 subtree=c("ENSG00000113494","ENSMODG00000020376"),
                 axis.text.size=5,
                 fontsize=7)
@ 
\caption{\genetreeexplanation{prolactin receptor, \textit{PRLR}}}
\label{fig:prolactin_receptor_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HOXA11" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 subtree=c("ENSBTAG00000014738","ENSECAG00000013229"),
                 axis.text.size=5,
                 fontsize=5
                 )
@ 
\caption{\genetreeexplanation{\textit{HOXA11}}}
\label{fig:hoxa11_evolution_tree}
\end{figure}

@@

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HBE1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=5)
@ 
\caption{\genetreeexplanation{\textit{HBE1}}}
\label{fig:hbe1_evolution_tree}
\end{figure}

@@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE) {
    temp.table <- 
        head(placenta.lineage.specific.analysis[analysis.name==analysis &
                                                p.value <= 0.05][order(p.value),
                                                                 list(human_name,p.value,fdr,fc),],
             n=5)
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[order(`p (t-test)`),]
        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}
output.top.genes.table <- function(data,output=TRUE) {
    if (output) {
        table.display <- c("s","s","s","d")
        table.align <- c("c","c","p{3in}","r")
        if (any(colnames(data) %in% "SEM"))  {
            table.display <- c(table.display,"d")
            table.align <- c(table.align,"r")
        }
        if (any(colnames(data) %in% "Human Symbol"))  {
            table.align <- c("c",table.align)
            table.display <- c("s",table.display)
        }
        print(xtable(data,
                     display=table.display,
                     align=table.align
                     ),
              booktabs=TRUE,
              floating=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)
    }
}

\begin{table}
  \footnotesize
<<rodentia.lineage.specific,echo=FALSE,results="asis">>=
rod.all.lin.spec <- placenta.lineage.specific.table("rod.all")
setkey(rod.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed in the Rodentia
  order as compared to all other analyzed species.}
\label{tab:rodentia_lineage_specific}
\end{table}

@@

\begin{table}
  \footnotesize
<<eua.lau.lineage.specific,echo=FALSE,results="asis">>=
eua.lau.lin.spec <- placenta.lineage.specific.table("eua.lau")
setkey(eua.lau.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed between the
  Euarchontoglires and Laurasiatherian superorder.}
\label{tab:laurasiatheria_euarchontoglires_lineage_specific}
\end{table}
